---
title: "3 Characterisation of PURA binding"
author: "Melina"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  BiocStyle::html_document:
    toc_float: TRUE
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{AS USP39 KD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(warning=FALSE, message=FALSE, cache=TRUE) #, fig.pos = "!H", out.extra = ""
```

# What is done here?

- Types of bound genes
- Distribution of binding sites over gene regions (with and without normalisation for region length)
- Distribution of crosslink event over gene regions, metagene profile 
- Distribution of binding sites per gene and in relation to TPM
- Relative maps of crosslink distribution
- GeneOntology and REACTOME enrichment analysis



```{r libraries, include=FALSE}
library(GenomicRanges)
library(rtracklayer)
library(knitr)
library(GenomicFeatures)
library(dplyr)
library(ggpubr)
library(purrr)
library(wesanderson)
library(cliProfiler)
library(hypeR)


report_color <- (pals::ocean.solar(15))

source("/Users/melinaklostermann/Documents/projects/PURA/02_R_new_pip/XX-helpful-chunks/theme_paper.R")
``` 

```{r}
################################################################################
###                       CODE REVIEW SUMMARY                                ###
################################################################################
### 
### Review by:          Mirko Brüggemann
### Comment shortcut:   MB
### Date:               23.09.2022
### Time spent:         ~2h

# Main objective: 
# Define binding profile for PURA binding sites, by overlapping with given 
# GENCODE annotation
# 
# Input:
# - PureCLIP binding sites from script 1
# - gene annotation as of gencode v31 as GFF3 and rds
# 
# Minor tasks the script performs are: 
# - match each binding site to a hosting transcript and transcript region
# - Collects additional features of binding sites (Strongest per gene, gene profile)
# - comparison with TPM values from a different dataset
# 
# Major/ issues:
# - Hard to follow the code logic, very little comments and variables are often not 
#   self explanatory. 
# - Some objects then do have explanatory names, but then they do not 
#   contain what the names suggests (eg strongest_BS does not contain binding sites,
#   but rather a sum of nts coverged by binding sites grouped by the region type).
# - In the part that identifies the bound gene type, the description said that the
#   longest transcript is used if multiple are present. But I found an example 
#   where this is not the case. (Maybe check this one, could also be that I got lost in the code)
# - The final output file (binding_sites_with_regions) contains a lot of NAs and 
#   is not clean (douplicated gene_id fields, that are sometimes NA and sometimes not)
#
# Minor/ suggestions:
# - The gene annotation is included once as gff3 and once as GRanges loaded from a 
#   precompiled rds file. And an additional txdb is generated and used as well.
#   This seems a bit redundant 
# - A lot of prints that produce unexpected result (eg. table(x) is used on object x
#   but the result is not a table, but rather a single number. )
# - Parts of the script contain code that is set to eval=FALSE, include=FALSE
# - At least one code chunk could not be evaluated since it took longer than 2 min
#   to execute. 
# 
# Nice job! 
# Feel free to review the review :) 
#
```

# Input

```{r input}


# Clip data
binding_sites <-  readRDS("/Users/melinaklostermann/Documents/projects/PURA/Molitor-et-al-2022/binding_sites.rds")  
CL_list <- readRDS("/Users/melinaklostermann/Documents/projects/PURA/Molitor-et-al-2022/bw_merges.rds")
#w_path_list <- readRDS(paste0("/Users/melinaklostermann/Documents/projects/PURA/02_R_new_pip/02-Comp-iCLIPs/XX-Combined_experiments/bw_path_list_PURA_all_iCLIPs.rds"))


# RNAseq
rnaseq_counts <- readRDS("/Users/melinaklostermann/Documents/projects/PURA/Molitor-et-al-2022/rnaseq_counts.rds")


# annotations
gff3 <- "/Users/melinaklostermann/Documents/projects/anno/GENCODEv31-p12/gencode.v31.annotation.gff3" # for cliProfiler
annotation <- readRDS("/Users/melinaklostermann/Documents/projects/PURA/Molitor-et-al-2022/annotation.rds")

### TODO MB
### For effciency, prebuild the TxDb and store it as .rds which is then only loaded upon execution.
### 
anno_txdb <- makeTxDbFromGRanges(annotation)

```


# Bound Gene types

```{r}
### TODO MB
### 1) when sorting GRanges, always sort seqlevels first -> sort(sortSeqlevels(x))
###

# anntotion of only genes, overlaps are resolved by keeping the longer gene
annotation_genes <- annotation[annotation$type=="gene"] 
annotation_genes <- sort(annotation_genes, by = ~ width + seqnames + start)

### TODO MB
### 1) unnecessary reassignment of bs = binding_sites
### 2) bs$idx is the same as bs$names
### 3) hard to follow code logic
### 4) not sure about the matching result
### -> in case of binding_sites_on_genes[120], you find 5 overlaps with annotated genes 
### -> subsetByOverlaps(annotation_genes, makeGRangesFromDataFrame(binding_sites_on_genes)[120])
### -> the binding site gets assigned to the gene with ID ENSG00000226396.1, but this is not the longest, but rather the shortest (please check ?)
###

#overlaps of crosslinks with genes
ol <- findOverlaps(annotation_genes, binding_sites)
bs <- binding_sites
bs$idx <- 1:NROW(bs)

# get all possible regions overlapping with clrosslinked nucleotide
bs_2 <- bs[subjectHits(ol)]
bs_2$gene_id <- annotation_genes[queryHits(ol)]$gene_id

# choose type highest in hiearcy
bs_3 <- as.data.frame(bs_2) %>%
  group_by(idx) %>%
  dplyr::slice(1)


binding_sites_on_genes <- bs_3  

# there are binding sites outside of annotated genes which are discarded here:
NROW(binding_sites)
NROW(binding_sites_on_genes)


bound_genes <- annotation_genes[substr(annotation_genes$gene_id,1,15) %in% substr(binding_sites_on_genes$gene_id,1,15)]  %>% # get bound genes
 # remove overlapping gene annotation
  as.data.frame(.)  %>%
  filter( !duplicated(gene_id) ) %>%
  filter( gene_type != "TEC" & !grepl(pattern = "pseudo", gene_type)  )# remove uninteresting types


### TODO MB
### plot works fine, but some suggestions here
### 1) use round(x, digits = 2) function for percentage labels
### 2) Sort by frequency
### 3) transform in log2 space

ggplot(bound_genes, aes(x = gene_type))+
  geom_bar(fill = "blue")+
  stat_count(aes(label=paste0(sprintf("%1.11f", ..count../sum(..count..)*100),
                              "% \n", ..count..), y=0.5*max(..count..)), 
             geom="text", colour="black", size=4, position=position_dodge(width=1)) +
  coord_flip()+
  theme_paper()

```



# Bound gene regions of protein coding genes

```{r}
### TODO MB
### does work as expected
###

# classify only protein coding genes, regions only from bound genes
anno_regions <- annotation[annotation$type %in% c("three_prime_UTR", "five_prime_UTR", "CDS") ]
anno_regions <- anno_regions[anno_regions$gene_id %in% binding_sites_on_genes$gene_id]
anno_regions <- anno_regions[anno_regions$gene_type == "protein_coding"]

annotation_bound <- annotation[annotation$gene_id %in% binding_sites_on_genes$gene_id]
annotation_bound_txdb <- makeTxDbFromGRanges(annotation_bound)

#get introns
introns_bound_genes <- intronsByTranscript(annotation_bound_txdb) %>% unlist(.)
introns_bound_genes$type = "intron"
anno_regions <- c(anno_regions, introns_bound_genes)

#overlaps of binding sites with genes
binding_sites_on_genes <- makeGRangesFromDataFrame(binding_sites_on_genes, keep.extra.columns = T)
ol_bs <- findOverlaps(anno_regions, binding_sites_on_genes, ignore.strand = FALSE, type ="any")
bs <- binding_sites_on_genes
bs$idx <- 1:NROW(binding_sites_on_genes)

# get all possible regions overlapping with clrosslinked nucleotide
bs_2 <- bs[subjectHits(ol_bs)]
elementMetadata(bs_2) <- c(elementMetadata(bs[subjectHits(ol_bs)]), elementMetadata(anno_regions[queryHits(ol_bs)]))

# choose type highest in hiearcy
bs_3 <- as.data.frame(bs_2) %>%
  group_by(idx) %>%
  arrange(factor(type, levels = c("three_prime_UTR", "five_prime_UTR", "CDS", "intron")), .by_group = T) %>%
  dplyr::slice(1)


ggplot(bs_3, aes(x=type))+
  geom_bar()+
  stat_count(aes(label=paste0(sprintf("%1.1f", ..count../sum(..count..)*100),
                              "% \n", ..count..), y=0.5*..count..), 
             geom="text", colour="white", size=4, position=position_dodge(width=1)) +
  theme_paper()+
  ylab("Number of crosslink events")


bs3 <- makeGRangesFromDataFrame(bs_3, keep.extra.columns = T)
bs_without_region <- binding_sites_on_genes[-queryHits(findOverlaps(binding_sites_on_genes, bs3))]
bs_without_region$type <- "no_region"

binding_sites_with_regions <- c(makeGRangesFromDataFrame(bs_3, keep.extra.columns = T) , bs_without_region)

```


```{r eval=FALSE, include=FALSE}

# make a GRanges with all regions overlaping BS
idx_BS_regions <- findOverlaps(binding_sites, anno_regions, ignore.strand = FALSE)
BS_with_all_regions <- binding_sites[queryHits(idx_BS_regions)]
elementMetadata(BS_with_all_regions) <- c(elementMetadata(BS_with_all_regions),
                                          elementMetadata(anno_regions[subjectHits(idx_BS_regions)]))

table(BS_with_all_regions$gene_type)

# hierarchy: 3'UTR > 5'UTR > CDS > intron 
# start and stop codon left out -> part of UTRs

# initialise column of peak region
BS_with_all_regions$BS_region <- NA

#########
# 3'UTR
#########
# get BS overlapping with 3'UTRs 
BS_3utr <- BS_with_all_regions[BS_with_all_regions$type == "three_prime_UTR"] %>% unique 
BS_3utr$BS_region <- "3UTR"
BS_assigned_regions <- BS_3utr


##########
# 5'UTR
#########
# get CDS BS that are not overlaping to UTRs
BS_5utr <- BS_with_all_regions[-queryHits(findOverlaps(
  BS_with_all_regions, BS_assigned_regions))] %>% .[.$type=="five_prime_UTR"]%>% unique
BS_5utr$BS_region <- "5UTR"
# add assigned CDS BS to assigned regions
BS_assigned_regions <- c(BS_assigned_regions, BS_5utr)

##########
# CDS
########
# get CDS BS that are not overlaping to UTRs
BS_cds <- BS_with_all_regions[-queryHits(findOverlaps(
  BS_with_all_regions, BS_assigned_regions))] %>% .[.$type=="CDS"]%>% unique
BS_cds$BS_region <- "CDS"
# add assigned CDS BS to assigned regions
BS_assigned_regions <- c(BS_assigned_regions, BS_cds)


NROW(BS_3utr)
NROW(BS_5utr)
NROW(BS_cds)

```

```{r eval=FALSE, include=FALSE}

# make a df with needed info from UTRs
utr_df <- as.data.frame(annotation[grepl(pattern = "UTR", annotation$type)])
#get cds and match start of cds to utr df
cds4UTR <- annotation[annotation$type=="CDS"]%>% as.data.frame
idx_UTR_cds <- match(utr_df$transcript_id, cds4UTR$transcript_id)
utr_df <- cbind(utr_df, start_CDS=cds4UTR[idx_UTR_cds,]$start)

# UTR is 3' if the start of the CDS is downstream of the start of the UTR, else is 5'
utr_df <- utr_df %>% 
  mutate(type_assigned=ifelse((.$strand == '+'& .$start > .$start_CDS)|
                       (.$strand=='-' & .$start <.$start_CDS), "3UTR", "5UTR"))
utr_35 <- utr_df %>% makeGRangesFromDataFrame(keep.extra.columns = TRUE)

table(utr_35$type_assigned)

table(anno_regions$type)

NROW(annotation[grepl(pattern = "UTR", annotation$type)])

```

```{r eval=FALSE, include=FALSE}
gff3 <- import.gff3("/Users/melinaklostermann/Documents/projects/anno/GENCODEv31-p12/gencode.v31.annotation.gff3") 
gtf <- import("/Users/melinaklostermann/Documents/projects/anno/GENCODEv31-p12/gencode.v31.annotation.gtf")

NROW(gff3[gff3$type=="three_prime_UTR"]) + NROW(gff3[gff3$type=="five_prime_UTR"])

NROW(gtf[gtf$type=="UTR"])


gff3[gff3$type %in% c("three_prime_UTR", "five_prime_UTR", "stop_codon")]
gtf[gtf$type=="UTR"]

gff3_txdb <- makeTxDbFromGRanges(gff3) 
gtf_txdb <- makeTxDbFromGRanges(gtf)

threeUTRsByTranscript(gff3_txdb) %>% NROW(.)
threeUTRsByTranscript(gtf_txdb) %>% NROW(.)


gff3_without_NA <- gff3[!is.na(gff3$transcript_support_level)]
gtf_without_NA <- gtf[!is.na(gtf$transcript_support_level)]

gff3_txdb <- makeTxDbFromGRanges(gff3_without_NA) 
gtf_txdb <- makeTxDbFromGRanges(gtf_without_NA)

threeUTRsByTranscript(gff3_txdb) %>% NROW(.)
threeUTRsByTranscript(gtf_txdb) %>% NROW(.)

```

## Normlisation for region length

```{r}
# classify only protein coding genes
anno_for_norm <- annotation[annotation$gene_type == "protein_coding"]
#anno_for_norm <- anno_for_norm[anno_for_norm$gene_id %in% binding_sites_on_genes$gene_id]
annotation_pc_txdb <-  makeTxDbFromGRanges(anno_for_norm)

anno_for_norm <- anno_for_norm[anno_for_norm$type %in% c("three_prime_UTR", "five_prime_UTR", "CDS") ]

#get introns
introns_for_norm <- intronsByTranscript(annotation_pc_txdb) %>% unlist(.)
introns_for_norm$type = "intron"
anno_for_norm <- c(anno_for_norm, introns_for_norm)
anno_for_norm_r <- GenomicRanges::reduce(anno_for_norm)

utr3 <- anno_for_norm[anno_for_norm$type=="three_prime_UTR"]
anno_left <- GenomicRanges::setdiff(anno_for_norm_r, utr3, ignore.strand=FALSE)

utr5 <- GenomicRanges::intersect(anno_left, anno_for_norm[anno_for_norm$type=="five_prime_UTR"] )
anno_left <- GenomicRanges::setdiff(anno_left, utr5, ignore.strand=FALSE)

cds <- GenomicRanges::intersect(anno_left, anno_for_norm[anno_for_norm$type=="CDS"] )
anno_left <- GenomicRanges::setdiff(anno_left, cds, ignore.strand=FALSE)

intron <- GenomicRanges::intersect(anno_left, anno_for_norm[anno_for_norm$type=="intron"] )
anno_left <- GenomicRanges::setdiff(anno_left, intron, ignore.strand=FALSE)


norm_df <- data.frame(region = c("three_prime_UTR", "five_prime_UTR", "CDS", "intron"),
                      width = c(sum(width(utr3)), sum(width(utr5)), sum(width(cds)), sum(width(intron))))
  

norm_regions <- bs_3 %>% group_by(type) %>% summarize(nts_in_BS = sum(width)) %>% 
  left_join(norm_df, by = c(type = "region")) %>%
  mutate(norm_binding_per_reg_size = nts_in_BS / width)



ggplot(norm_regions, aes(x=type, y = norm_binding_per_reg_size))+
  geom_col()+
  theme_paper()

```

# Strongest binding site per gene
```{r}
### TODO MB
### here the pureclip score is used to define the strongest binding site per gene
### this is then scaled by the number of nts covered by binding site on that gene
### -> I don't really get why
### -> It's not shown in plot or in description
### 

strongest_BS <- bs_3 %>% group_by(gene_id) %>%
  arrange(desc(score)) %>%
  dplyr::slice(1) %>%
  ungroup() %>%
  group_by(type)%>%
  summarize(nts_in_BS = sum(width)) %>% 
  left_join(norm_df, by = c(type = "region")) %>%
  mutate(norm_binding_per_reg_size = nts_in_BS / width)



ggplot(strongest_BS, aes(x=type, y = norm_binding_per_reg_size))+
  geom_col()+
  theme_paper()

```


# Distribution of crosslinks on gene regions

```{r eval=FALSE, include=FALSE}
### TODO MB
### Code takes very long to execute
### -> Stopped execution after ~2min
anno_regions_all <- annotation[annotation$type %in% c("three_prime_UTR", "five_prime_UTR", "CDS") ]
#anno_regions_all <- anno_regions_all[anno_regions_all$gene_id %in% binding_sites_on_genes$gene_id]
anno_regions_all <- anno_regions_all[anno_regions_all$gene_type == "protein_coding"]

annotation_txdb <- makeTxDbFromGRanges(annotation_bound)

#get introns
introns_bound_genes <- intronsByTranscript(annotation_txdb) %>% unlist(.)
introns_bound_genes$type = "intron"
anno_regions_all <- c(anno_regions_all, introns_bound_genes)

#overlaps of binding sites with genes
cl <- c(CL_list[[1]], CL_list[[2]])
ol_cl <- findOverlaps(anno_regions_all, cl, ignore.strand = FALSE, type ="any")
cl$idx <- 1:NROW(cl)

# get all possible regions overlapping with clrosslinked nucleotide
cl_2 <- cl[subjectHits(ol_cl)]
elementMetadata(cl_2) <- c(elementMetadata(cl[subjectHits(ol_cl)]), elementMetadata(anno_regions_all[queryHits(ol_cl)]))

# choose type highest in hiearcy
cl_3 <- as.data.frame(cl_2) %>%
  group_by(idx) %>%
  arrange(factor(type, levels = c("three_prime_UTR", "five_prime_UTR", "CDS", "intron")), .by_group = T) %>%
  dplyr::slice(1)


ggplot(cl_3, aes(x=type))+
  geom_bar()+
  stat_count(aes(label=paste0(sprintf("%1.1f", ..count../sum(..count..)*100),
                              "% \n", ..count..), y=0.5*..count..), 
             geom="text", colour="white", size=4, position=position_dodge(width=1)) +
  theme_paper()+
  ylab("Number of crosslink events")

#ggsave(paste0(outpath, "crosslink_gene_regions.pdf"), width = 5 , height  = 5, units = "cm")

```


# Meta gene profile

```{r eval=FALSE, include=FALSE}
# split for crosslinks on genes with highestbbinding in cds vs utr
BS_cds_utr <- binding_sites_with_regions[binding_sites_with_regions$type %in% c("CDS", "three_prime_UTR")]
highest_BS_per_gene <- BS_cds_utr %>%
  as.data.frame(.) %>%
  group_by(gene_id) %>%
  arrange(desc(score), .by_group = T) %>%
  dplyr::slice(1) %>%
  ungroup() %>%
  mutate(gene_id = substr(gene_id,1,15))

# selct only crosslinks on annotated genes
idx_cl_gene_id <- findOverlaps(cl, annotation[annotation$type == "gene"])
idx_cl_gene_id <- as.data.frame(idx_cl_gene_id) %>%  filter(!duplicated(queryHits))

cl$gene_id <- NA
anno_gene <- annotation[annotation$type == "gene"]
cl[idx_cl_gene_id$queryHits]$gene_id <- substr(anno_gene[idx_cl_gene_id$subjectHits]$gene_id,1,15)

cl_on_genes <- cl[!is.na(cl$gene_id)] %>%
  as.data.frame(.) %>%
  left_join(highest_BS_per_gene[,c("gene_id", "type")], by = "gene_id") %>% 
  filter(!is.na(type))%>%
  mutate(group = type) %>%
  makeGRangesFromDataFrame(keep.extra.columns = T)

# make metaprofile 

profile <- metaGeneProfile(cl_on_genes, annotation = gff3, group="type")
profile$Plot + ylim(c(0,0.9)) + theme_paper() + theme(legend.position = NULL)

#ggsave(paste0(outpath,"MetaGeneProfile_cds_utr.pdf" ), width = 9, height = 5, units = "cm")

```

# Binding vs tpm

```{r}
### TODO MB
### tpm cutoffs are not known
### 

s <- list(0,1,10,100,1000)

RNA_tpm_cutoffs <- map(s, ~rnaseq_counts %>% filter(mean_tpm_rnaseq > .x))

RNA_tpm_cutoffs_bound <- map2_dfr(RNA_tpm_cutoffs, s, ~c(bound = nrow(.x[.x$gene_id %in% substr(binding_sites_with_regions$gene_id,1,15),]),
                                                  all_over_tpm_cut =   nrow(.x),
                                                  tpm_cut = .y))


RNA_tpm_cutoffs_bound <- mutate(RNA_tpm_cutoffs_bound, ratio_bound = bound / all_over_tpm_cut)

ggplot(RNA_tpm_cutoffs_bound, aes(x = as.factor(tpm_cut), y = ratio_bound, label = all_over_tpm_cut))+
  geom_col(position = "identity")+
  geom_text(aes(label=paste(round(ratio_bound,2), "% \n of", all_over_tpm_cut)), 
             geom="text", color= "black", size=4)+
  #ggrepel::geom_text_repel()+
  ylim(0,1)+
  theme_paper()



```

## RNAseq TPM vs CLIP TPM

```{r}
#########################
# Calcualte CLIP tpms
#########################

# get transcript lengths
transcript_length <- transcriptLengths(anno_txdb, with.cds_len = T)
transcript_length_mean <- transcript_length %>%
  dplyr::group_by(gene_id) %>%
  summarise(transcript_length = mean(tx_len), .groups = "keep")
transcript_length_mean$gene_id <- substr(transcript_length_mean$gene_id,1,15)
binding_sites_with_regions$gene_id <- substr(binding_sites_with_regions$gene_id,1,15)

# add mean transcript length to binding sites
binding_sites_with_regions <- binding_sites_with_regions %>% as.data.frame() %>%
  left_join(., transcript_length_mean, by = "gene_id", ) 

# add gene length for genes with no annotated transcript
binding_sites_with_regions[is.na(binding_sites_with_regions$transcript_length),]$binding_sites_with_regions <-  binding_sites_with_regions[is.na(binding_sites_with_regions$transcript_length),]$width

# tpm for all samples
binding_sites_with_regions <- binding_sites_with_regions %>% mutate(
    s1.clip.tpm = clp_rep1 / transcript_length,
    s2.clip.tpm = clp_rep2 / transcript_length,
    s3.clip.tpm = clp_rep3 / transcript_length,
    s4.clip.tpm = clp_rep4 / transcript_length
 )


# scaling factor values for samples 
scaling_factor <- binding_sites_with_regions[,c("s1.clip.tpm", "s2.clip.tpm", "s3.clip.tpm", "s4.clip.tpm")] 
scaling_factor <- colSums(scaling_factor, na.rm = T)
scaling_factor <- scaling_factor/1000000

binding_sites_with_regions <- binding_sites_with_regions %>% mutate(
    s1.clip.tpm = s1.clip.tpm / scaling_factor["s1.clip.tpm"],
    s2.clip.tpm = s2.clip.tpm / scaling_factor["s2.clip.tpm"],
    s3.clip.tpm = s3.clip.tpm / scaling_factor["s3.clip.tpm"],
    s4.clip.tpm = s4.clip.tpm  / scaling_factor["s4.clip.tpm"]) %>%
  rowwise() %>%
  mutate( mean.clip.tpm = mean(c(s1.clip.tpm, s2.clip.tpm, s3.clip.tpm, s4.clip.tpm))
    )


#############################
# add RNAseq TPM and compare
############################
binding_sites_with_regions <- binding_sites_with_regions %>% left_join(., rnaseq_counts, by = c(gene_id= "gene_id"))

ggplot(binding_sites_with_regions, aes(x = log10(mean_tpm_rnaseq), log10(mean.clip.tpm)))+
                ggrastr::rasterise(geom_point(size = 0.2), dpi = 300)+
  ggrastr::rasterise(ggpointdensity::geom_pointdensity(size = 0.2), dpi = 300)+
  geom_smooth(method = "lm", color = "black")+
  stat_cor()+
  theme_paper()
  
#ggsave(filename = paste0(output_path, "/tpm_comp.pdf"), width = 7, height = 5, units = "cm")


```


# Number of binding sites per gene

```{r}
binding_sites_with_regions <- makeGRangesFromDataFrame(binding_sites_with_regions, keep.extra.columns = T)
bound_genes <- makeGRangesFromDataFrame(bound_genes, keep.extra.columns = T)

n_BS_per_gene <- data.frame( n_BS = as.numeric(countOverlaps(bound_genes, binding_sites_with_regions)),
                             gene_id = substr(bound_genes$gene_id,1,15)) %>%
  mutate(., n_BS_plot = case_when (n_BS > 100 ~ 100,
                                   T ~ n_BS))


ggplot(n_BS_per_gene, aes(x=n_BS_plot))+
  geom_histogram(binwidth = 1)+
  theme_paper()+
  geom_vline(xintercept = mean(n_BS_per_gene$n_BS), color =  "blue")+
  geom_vline(xintercept = median(n_BS_per_gene$n_BS), color =  "green")+
  geom_vline(xintercept = quantile(n_BS_per_gene$n_BS)[2], color = "green")+
   geom_vline(xintercept = quantile(n_BS_per_gene$n_BS)[4], color =  "green")+
  xlab("Number of binding sites per gene")+
  ylab("Number of genes")


```



# Number of binding sites
From revision:
 - the distribution of the number of sites per gene over the whole transcriptome should be shown. Is it a continuum from no binding sites to many, or is there a subset of genes intensively bound
```{r}
###################
# Number of BS per gene
##################

n_BS_per_gene_tpm <- left_join(n_BS_per_gene, rnaseq_counts, by = "gene_id" )

### TODO MB
### this should probably be viewed on log2 scale
###

ggplot(n_BS_per_gene_tpm, aes(x=mean_tpm_rnaseq, y = n_BS))+
  geom_point()+
  #ggforce::facet_zoom(xlim=c(0,1500))+
  #geom_hline(yintercept = 25, color ="red")+
  geom_smooth()+
  ylab("Number of binding sites per gene")+
  xlab("Gene expression [tpm]")+
  theme_paper()



n_BS_per_gene_tpm <- n_BS_per_gene_tpm %>% mutate(bin = cut(mean_tpm_rnaseq, breaks = c(-Inf,1,10,100,200,300,400,500,600,25000 )) )

median_per_bin <- n_BS_per_gene_tpm %>% group_by(bin) %>%
  summarise(median = median(n_BS))

ggplot(n_BS_per_gene_tpm, aes(x=as.factor(bin), y = n_BS))+
  # gghalves::geom_half_boxplot(outlier.size = -5, lwd = 0.5) +
  ggrastr::rasterize(geom_jitter(), dpi = 400)+
  geom_point(data = median_per_bin, aes(x = bin, y = median ), color = "red")+
  theme_paper()+
  ylab("Number of binding sites per gene")+
  xlab("Binned gene expression [tpm]")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#ggsave(paste0(outpath,"tpm_vs_BS_per_gene.pdf"), height = 6, width = 7, units = "cm")

```


# GO

```{r}
# background: all genes with crosslinks
expr_genes <- subsetByOverlaps(annotation_genes, c(CL_list[[1]],CL_list[[2]]))

# msigdb_available(species = "Homo sapiens")
reactome_geneset <- msigdb_gsets("Homo sapiens", "C2", "CP:REACTOME")
cell_comp_geneset <- msigdb_gsets("Homo sapiens", "C5", "CC")


# background
bg <- unique(expr_genes$gene_name)

# Test USS groups vs gene sets
hyps_reactome <-  hypeR(bound_genes$gene_name, reactome_geneset, test="hypergeometric", fdr=1, pval = 1, background = bg )
hyps_cell_comp <- hypeR(bound_genes$gene_name, cell_comp_geneset, test="hypergeometric", fdr=1, pval = 1, background = bg) 



# plot reactome
reactome_df <- hyps_reactome$data %>% mutate(label = tolower(label), generatio = overlap / geneset)

ggplot(reactome_df[1:25,], aes(y = generatio, x = factor(label, level = rev(unique(label)))))+
  geom_col()+
  coord_flip()+
  theme_paper()+
  scale_size(range = c(1, 3))

# xlsx::write.xlsx(reactome_df, paste0(outpath, "iCLIP_reactome_list.xlsx"))
# ggsave(paste0(outpath, Sys.Date(), "reactome_iCLIP_paper.pdf"), width = 17, height = 8, units = "cm")


# plot cellular components 
cc_df <- hyps_cell_comp$data %>% mutate(label = tolower(label), generatio = overlap / geneset)

ggplot(cc_df[1:25,], aes(y = generatio, x = factor(label, level = rev(unique(label)))))+
  geom_col()+
  coord_flip()+
  theme_paper()+
  scale_size(range = c(1, 3))

#xlsx::write.xlsx(cc_df, paste0(outpath, "iCLIP_GO_cc_list.xlsx"))
#ggsave(paste0(outpath, Sys.Date(), "cc_iCLIP_paper.pdf"), width = 17, height = 8, units = "cm")


# cellular components connected to neurons
dendritic_cc <- cc_df[grepl(cc_df$label, pattern="dendritic"),]
neuron_cc <- cc_df[grepl(cc_df$label, pattern="neuron"),]
axon_cc <- cc_df[grepl(cc_df$label, pattern="axon"),]

neuro_dendritic_cc <- rbind(dendritic_cc, neuron_cc, axon_cc)

ggplot(neuro_dendritic_cc[neuro_dendritic_cc$fdr < 0.05,], aes(y = generatio, x = factor(label, level = rev(unique(label))), fill = -log10(fdr) ))+
  geom_col()+
  coord_flip()+
  theme_paper()+
  scale_size(range = c(1, 3))

#ggsave(paste0(outpath, Sys.Date(), "cc_dndritic_iCLIP_paper.pdf"), width = 17, height = 8, units = "cm")


```


# Output
- binding sites with additional infos (binding_sites_characterized.rds)

```{r eval=FALSE, include=FALSE}

### TODO MB
### final object contains a lot of NAs
### in some columns it's expected, in others not
### eg. the gene_id field seems to be duplicated (gene_id.1), but it's not always filled? 

saveRDS(binding_sites_with_regions,  "/Users/melinaklostermann/Documents/projects/PURA/Molitor-et-al-2022/binding_sites_characterized.rds")

```


