---
title: "4 Comparison of the crosslink patterns in three different PURA iCLIP experiments from HeLa cells and NPC PURA iCLIP"
author: "Melina"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(warning=FALSE, message=FALSE, cache=TRUE) #, fig.pos = "!H", out.extra = ""
```

```{r libraries, include=FALSE}
library(GenomicRanges)
library(rtracklayer)
library(knitr)
library(GenomicFeatures)
library(dplyr)
library(ggpubr)
library(Gviz)
library(purrr)
library(wesanderson)
library(ComplexHeatmap)
library(BSgenome.Hsapiens.UCSC.hg38)


report_color <- (pals::ocean.solar(15))

outpath <- "/Users/melinaklostermann/Documents/projects/PURA/02_R_new_pip/01-BS_def/02-BS_def_endo/Report59-output/"
source("/Users/melinaklostermann/Documents/projects/PURA/02_R_new_pip/XX-helpful-chunks/theme_paper.R")
```

```{r}
################################################################################
###                       CODE REVIEW SUMMARY                                ###
################################################################################
### 
### Review by:          Mirko Brüggemann
### Comment shortcut:   MB
### Date:               04.10.2022
### Time spent:         ~2h
# Main objective: 
# Comparison of different PURA iCLIP datasets, mainly binding in NPC cells
# 
# Input:
# - PURA crosslinks from the preprocessing as bw
# - gene annotation as of gencode v31 as rds
# - Binding sites from script 1
# 
# Minor tasks the script performs are: 
# - 
# 
# Major/ issues:
# - No major issues, code works as expected :) 
#
# Minor/ suggestions:
# - More comments in the code this time which is nice :)
# - Still the code is often hard to read
#   -> multiple reassignments of values between variables within the same code junk
#   -> input data is loaded in partly redundant lists 
#   -> a lot of hard coded values 
# - GVIZ functions are slow because iCLIP signal is loaded as GRanges instead of RLE
# - calculation of iCLIP coverage over a GRanges object can be optimized by using BindingSiteFinder
# 
# Nice job! 
# Feel free to review the review :) 
#
```


# What is done here?

- Metaprofile of PURA binding in NPC cells (iCLIP experiment with low signal)
- Some GVIZ plots of PUAR crosslinking patterns in HeLa snd NPC cells
- Min max normalised spline smoothed heatmaps of crosslinks from endogenous PURA in Hela, and FLAG-PURA overexpression pulled down with two different antibodies (antiFALG & antiPURA)

# Input

```{r}
# load BS with overlap annotation
BS <- readRDS("/Users/melinaklostermann/Documents/projects/PURA/Molitor-et-al-2022/binding_sites_characterized.rds")

# crosslinks from NPC cells
npc_cl_p_rle <- import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/PURA_NPC_and_FLAG/imb_koenig_2020_17_koenig_iCLIP_PURA/merged/bw/imb_koenig_2020_17_NPC.v2uniqMD.duprm.plus.bw", as = "Rle")
npc_cl_m_rle <- import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/PURA_NPC_and_FLAG/imb_koenig_2020_17_koenig_iCLIP_PURA/merged/bw/imb_koenig_2020_17_NPC.v2uniqMD.duprm.minus.bw",  as = "Rle")

npc_cl_p_gr <- import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/PURA_NPC_and_FLAG/imb_koenig_2020_17_koenig_iCLIP_PURA/merged/bw/imb_koenig_2020_17_NPC.v2uniqMD.duprm.plus.bw")
npc_cl_m_gr <- import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/PURA_NPC_and_FLAG/imb_koenig_2020_17_koenig_iCLIP_PURA/merged/bw/imb_koenig_2020_17_NPC.v2uniqMD.duprm.minus.bw")

# crosslinks from all three HeLa iCLIPs
hela_cl_rle <- list(endo_p = import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/PURA_endo/imb_koenig_2020_07_koenig_iCLIP_PURA_endogene/merged/bw/imb_koenig_2020_07_PURAendo.v2uniqMD.duprm.plus.bw", as = "Rle"),
                    endo_m = import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/PURA_endo/imb_koenig_2020_07_koenig_iCLIP_PURA_endogene/merged/bw/imb_koenig_2020_07_PURAendo.v2uniqMD.duprm.minus.bw", as = "Rle"),
                    oe_pura_p = import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/oe-imb_koenig_2019_11_koenig_iCLIP_PURA___v2uniqMD/merged/bw/imb_koenig_2019_11_allsamples.v2uniqMD.duprm.plus.bw", as = "Rle"),
                    oe_pura_m = import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/oe-imb_koenig_2019_11_koenig_iCLIP_PURA___v2uniqMD/merged/bw/imb_koenig_2019_11_allsamples.v2uniqMD.duprm.minus.bw", as = "Rle"),
                    oe_flag_p = import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/PURA_NPC_and_FLAG/imb_koenig_2020_17_koenig_iCLIP_PURA/merged/bw/imb_koenig_2020_17_flag.v2uniqMD.duprm.plus.bw", as = "Rle"),
                    oe_flag_m = import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/PURA_NPC_and_FLAG/imb_koenig_2020_17_koenig_iCLIP_PURA/merged/bw/imb_koenig_2020_17_flag.v2uniqMD.duprm.minus.bw", as = "Rle"))

### TODO MB
### this is redundant to the hela_cl_rle list
###
# endo HeLa crosslinks for GVIZ plots
hela_cl <- list(endo_p = import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/PURA_endo/imb_koenig_2020_07_koenig_iCLIP_PURA_endogene/merged/bw/imb_koenig_2020_07_PURAendo.v2uniqMD.duprm.plus.bw"),
                endo_m = import.bw("/Users/melinaklostermann/Documents/projects/PURA/01_raw_data/PURA_endo/imb_koenig_2020_07_koenig_iCLIP_PURA_endogene/merged/bw/imb_koenig_2020_07_PURAendo.v2uniqMD.duprm.minus.bw"))

### TODO MB
### would be faster if you load the annotation with loadDb() instead of doing it again from scratch
###
# filtered annotation
annotation <- readRDS("/Users/melinaklostermann/Documents/projects/PURA/Molitor-et-al-2022/annotation.rds")
anno_txdb <- makeTxDbFromGRanges(annotation)


# setting for gviz plots 
gen <- "hg38"
cex <- 0.7
cex.a <- 0.5
```


# Meta Profile NPC

```{r meta_profile_npc, fig.height=5, fig.width=5}
#############################
# metaprofile of npc crosslinks in 15nt window on PURA BS
###########################

# enlarge BS
BS_15nt <- as.data.frame(BS+10 )%>%
  arrange(desc(score)) %>%
  .[1:floor(NROW(BS)),] %>%
  makeGRangesFromDataFrame(keep.extra.columns = T)

npc_cl_p <- as.matrix(npc_cl_p_rle[BS_15nt[strand(BS_15nt) == "+"]])
npc_cl_m <- as.matrix(npc_cl_m_rle[BS_15nt[strand(BS_15nt) == "-"]])
npc_cl <- rbind(npc_cl_p, npc_cl_m) 

# exclude windows with 0 NPC crosslinks
 npc_cl_without_0 <- npc_cl[rowSums(npc_cl)!=0 ,] %>% # %>% 
   as.data.frame()%>%
   arrange(desc(rowSums(.))) %>%
   as.matrix()

# make heatmap
col_fun <-  circlize::colorRamp2(c(0,10), c("white", "black"))
line <- HeatmapAnnotation(crosslinks = anno_lines(colSums(npc_cl), height = unit(2, "cm"), ylim = c(0, 13000)))

ht_opt$heatmap_column_names_gp = gpar(fontsize = 5)

### TODO MB
### here i was a bit confused by the heatmap itself, not the code.
### The colsums shows the sum and goes to up to 13000, but the heatmap color scale 
### goes to a max of 10. I can't see how the displayed values would add up the way 
### that they are color coded atm
### -> also adjust box limitis to avoid that line reaches out of box
###
Heatmap(npc_cl_without_0[1:1000,], col =col_fun, cluster_columns = F, cluster_rows = F, top_annotation = line, use_raster = T, raster_device = "png", raster_quality = 10)


### TODO MB
### this can be added directly using linkAnnotations from ComplexHeatmaps
###
# names afterwards added in affinity designer
npc_cl_top_genes <- BS_15nt$gene_name
npc_cl_top_genes[duplicated(npc_cl_top_genes)] <- ""

npc_cl_top_genes

```

# GVIZ examples Hela endogenous PURA and NPC endogenous PURA crosslinking

```{r}
############################
# function for GVIZ plots
############################

### TODO MB
### takes very long to execute, maybe use system.time({}) to track the individual functions parts for speedup
### Plotting could be done in a loop and piped into a pdf document directly
###
plotGviz <- function(GR, i, w,  cl_plus, cl_minus, name, anno, BS, color){
  # get chromosome amd range, make data frame
  chr<- as.character(seqnames(GR[i]))
  df <- GR[i,] %>% as.data.frame
  range <- GRanges(seqnames=chr, strand=strand(GR[i]),  ranges=IRanges(start= start(GR[i])-w[1], end = end(GR[i])+w[2])) 
  
  # make track of crosslinks depending on which strand
  # cl_2
  if(df$strand=="+"){
    cl_at_region <- subsetByOverlaps(cl_plus, range) %>% keepStandardChromosomes()
    dTrack <-
      DataTrack(
        cl_at_region,
        genome = gen,
        name = "cl endo",
        type = "histogram", cex.title=cex, cex.axis= cex.a, background.title = "white", col.axis= "black", fontcolor = "black")
  }else{
    cl_at_region <- subsetByOverlaps(cl_minus, range) %>% keepStandardChromosomes()
    dTrack <-
      DataTrack(
        cl_at_region,
        genome = gen,
        name = "cl_endo",
        type = "histogram", cex.title=cex,  cex.axis= cex.a, background.title = "white", col.axis= "black", fontcolor = "black")
  }
  
 if(df$strand=="+"){ 
    seqTrack <- SequenceTrack(Hsapiens, chromosome=chr, complement = F)
  }else{
    seqTrack <-SequenceTrack(Hsapiens, chromosome=chr, complement = T)} 
  
  #pureclip track
  BS_range <- BS[strand(BS)==strand(GR[i])] %>% subsetByOverlaps(.,GR[i]+w[1])

  ### TODO MB
  ### report_color is global 
  ###
  BS_new_track <- GeneRegionTrack(BS_range, genome = gen, name="BS endo",cex.axis= cex.a,
                                  col.histogram=report_color[13], fill=report_color[5], stacking="hide", rotation.title=0, cex.title=cex, background.title = "white", fontcolor = "black")
  
  
  # track showing loaction
  gtrack <- GenomeAxisTrack(cex.title= 2, add53 = T)
  # annotation track
  anno_track <- GeneRegionTrack(anno, genome = gen, chromosome = chr,
                                name = "anno", fill = color, cex.title=cex, cex.axis= cex.a, stacking = "squish", background.title = "transparent", fontcolor = "black")
  
  plotTracks(list(dTrack, seqTrack, BS_new_track, anno_track, gtrack),
               from = start(range), to = end(range),
               chromosome=chr, sizes =c(5,1,1,1.5, 1), 
               main= paste(GR[i]$gene_name, "-", chr, ":", start(range),"-", end(range)), cex.main = 1)
}

##########################
# make GVIZ plots
########################

#=================================
# Highest crosslinked regions in STARD7, SEC61A1, YWHAE  zoom-in endo. PURA HeLa
#==================================

regions_for_gviz_1 <- BS %>% 
  as.data.frame(.) %>%
  group_by(gene_id) %>%
  arrange(desc(score), .by_group = T) %>% 
  dplyr::slice(1) %>% 
  ungroup %>%
  filter(gene_name %in% c("STARD7", "SEC61A1", "YWHAE")) %>%
  makeGRangesFromDataFrame(keep.extra.columns = T)

# SEC61A1
plotGviz(GR = regions_for_gviz_1[1]  , i = 1, w = c(150,150),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_1[1]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

plotGviz(GR = regions_for_gviz_1[1]  , i = 1, w = c(150,150),  cl_plus = npc_cl_p_gr, cl_minus = npc_cl_m_gr, name = regions_for_gviz_1[1]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

# STARD7
plotGviz(GR = regions_for_gviz_1[2]  , i = 1, w = c(150,170),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_1[2]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

plotGviz(GR = regions_for_gviz_1[2]  , i = 1, w = c(150,170),  cl_plus = npc_cl_p_gr, cl_minus = npc_cl_m_gr, name = regions_for_gviz_1[1]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

# STARD7 whole 3'UTR
plotGviz(GR = regions_for_gviz_1[2]  , i = 1, w = c(300,2000),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_1[2]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

# YWHAE
plotGviz(GR = regions_for_gviz_1[3]  , i = 1, w = c(150,170),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_1[3]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

plotGviz(GR = regions_for_gviz_1[3]  , i = 1, w = c(150,170), cl_plus = npc_cl_p_gr, cl_minus = npc_cl_m_gr, name = regions_for_gviz_1[3]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")


#=================================
# non-coding examples RN7SL1, MALAT1, NEAT1 
#==================================
regions_for_gviz_2 <- BS %>% 
  as.data.frame(.) %>%
  group_by(gene_id) %>%
  arrange(desc(score), .by_group = T) %>% 
  dplyr::slice(1) %>% 
  ungroup %>%
  filter(gene_id %in% c("ENSG00000251562", "ENSG00000245532", "ENSG00000276168" )) %>% # MALAT1
  makeGRangesFromDataFrame(keep.extra.columns = T)

# MALAT1
plotGviz(GR = regions_for_gviz_2[2]  , i = 1, w = c(400,400),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_1[2]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

# NEAT1
plotGviz(GR = regions_for_gviz_2[1]  , i = 1, w = c(400,400),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_1[1]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

# RN7SL1
plotGviz(GR = regions_for_gviz_2[3]  , i = 1, w = c(200,100),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_1[3]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")


#=================================
# RNA of p-body essential proteins: LSM14A, DDX6 (whole gene)
#==================================
regions_for_gviz_3 <- annotation[annotation$type == "gene"] %>% 
  as.data.frame(.) %>%
  filter(gene_name %in% c("LSM14A", "DDX6") ) %>%
  makeGRangesFromDataFrame(keep.extra.columns = T)

# LSM14A
plotGviz(GR = regions_for_gviz_3[2]  , i = 1, w = c(0,0),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_1[2]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

# DDX6
plotGviz(GR = regions_for_gviz_3[1]  , i = 1, w = c(0,0),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_1[1]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")




#=================================
# Bound regions tested in EMSAs
#==================================
# regions of seqeunces used for emsa
regions_for_gviz_4 <- GRanges(c("chr2:96185058-96185112:-", "chr11:65441947-65441998:+", "chr5:86619374-86619444:+", "chr19:34221504-34221561:+") , gene_name = c("STARD7", "NEAT1", "COX7C", "LSM14A" ))

# STARD7
plotGviz(GR = regions_for_gviz_4[1]  , i = 1, w = c(0,0),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_1[1]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

# NEAT1
plotGviz(GR = regions_for_gviz_4[2]  , i = 1, w = c(0,0),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_4[2]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

# SCOX7C
plotGviz(GR = regions_for_gviz_4[3]  , i = 1, w = c(0,0),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_4[3]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")

#  LSM14A
plotGviz(GR = regions_for_gviz_4[4]  , i = 1, w = c(0,0),  cl_plus =hela_cl[[1]], cl_minus = hela_cl[[2]], name = regions_for_gviz_4[4]$gene_name , anno = anno_txdb, BS = BS, color = "darkred")


```

# Compare three PURA HeLa CLIPs with spline smoothed maps

```{r eval=FALSE, include=TRUE}
### TODO MB
### why eval FALSE, there is nothing computed here
###
### ===============================================
###  Normalisation and spline smoothing of crosslink patterns
### ===============================================
# Note: this approach was adapted from Heyl & Backofen 2021 - "StoatyDive: Evaluation and classification of peak profiles for sequencing data"

# min max normalisation
min_max_noramlize <- function(x){
  x_new <- (x-min(x)) / (max(x)-min(x)) 
  return(x_new)
}


# smooth
smoothing <- function(y, lambda, dim){
  data_points <- length(y)
  
  # test if we have just a vector of constant values.
  which(y != max(y))
  if ( length(which(y != max(y))) != 0 ){
    x <- c(1:data_points)
    y.smooth <- smooth.spline(x, y, spar=lambda)
    x_new <- seq(1, data_points, data_points/dim)
    y_new <- predict(y.smooth, x_new)$y
    y_new <- y_new/max(y_new)
    return(y_new)
  }else{ 
    x_new <- seq(1, data_points, data_points/dim)
    y_new <- rep( max(y), length(x_new) )
    return(y_new)}
}


clean_smooth <- function(data_smoothed) {
  
  data_smoothed <- t(data_smoothed)
  # get rid of negative values
  data_smoothed[which(data_smoothed < 0)] = 0.0
  
  # get rid of diract impulses bigger than max (> 1.0)
  data_smoothed[which(data_smoothed > 1.0)] = 1.0 
  
  return(data_smoothed)
}
```

## in the beginning of 3'UTRs

```{r eval=FALSE, include=TRUE}
### TODO MB
### try to avoid re-assigning variables may times, this makes the code hard to understand
### calculating the coverage of iCLIP counts over a GRanges object can be done wtih BindingSiteFinder, 
### which is easier to read and faster
###

### ============================================
###  Crosslink patterns at the start of 3'UTRs
### ============================================

# window at the beginning of the 3'UTR
# select UTRs
three_UTRs <- annotation[annotation$type=="three_prime_UTR"] 
three_UTRs <- three_UTRs[!duplicated(three_UTRs$gene_id)]
w = 300

window <- three_UTRs
end(window[strand(window)=="+"]) <- start(window[strand(window)=="+"])+300
start(window[strand(window)=="-"]) <- end(window[strand(window)=="-"])-300


cl_window <- c(map(hela_cl_rle[c(1,3,5)], ~.x[window[strand(window)=="+"]] %>% as.matrix() ),
               map(hela_cl_rle[c(2,4,6)], ~.x[window[strand(window)=="-"]] %>% as.matrix() %>% .[,ncol(.):1] ))

# select 3'UTRs with medium coverage
cl_window_coverage_p <- map_dfc(cl_window[1:3], ~ rowSums(.x))
cl_window_coverage_m <- map_dfc(cl_window[4:6], ~ rowSums(.x))

cov_p <- map_dfc(cl_window_coverage_p, ~.x) %>% mutate(medium_cov =
                                                            case_when( endo_p > 100 &
                                                                        oe_pura_p > 100 &
                                                                        oe_flag_p > 100 &
                                                                        endo_p < 10^6 &
                                                                        oe_pura_p < 10^6 &
                                                                        oe_flag_p < 10^6 
                                                                      ~ T,
                                                                      T~F))
cov_m <- map_dfc(cl_window_coverage_m, ~.x) %>% mutate(medium_cov =
                                                            case_when( endo_m > 100 &
                                                                        oe_pura_m > 100 &
                                                                        oe_flag_m > 100 &
                                                                        endo_m < 10^6 &
                                                                        oe_pura_m < 10^6 &
                                                                        oe_flag_m < 10^6 
                                                                      ~ T,
                                                                      T~F))


cl_window <- c(map(cl_window[1:3], ~.x[cov_p$medium_cov,] ),
               map(cl_window[4:6], ~.x[cov_m$medium_cov,] ))


window_p <- window[strand(window)=="+"]
window_m <- window[strand(window)=="-"]
window <- c(window_p[cov_p$medium_cov], window_m[cov_m$medium_cov])


cl_window <- list(endo = rbind(cl_window[[2]], cl_window[[5]]),
                  oe  = rbind(cl_window[[1]], cl_window[[4]]),
                  flag = rbind(cl_window[[3]], cl_window[[6]]))


# normalise
cl_window_normalized <- lapply(cl_window, function(x) apply(x, 1, min_max_noramlize) %>% t(.))


# set NaNs to zero
cl_window_normalized[[1]][which(is.na(cl_window_normalized[[1]]))] = 0
cl_window_normalized[[2]][which(is.na(cl_window_normalized[[2]]))] = 0
cl_window_normalized[[3]][which(is.na(cl_window_normalized[[3]]))] = 0

# smoothing
cl_window_norm_smooth <- lapply(cl_window_normalized,  apply, 1, function(x) smoothing(y=x, lambda = 0.5, dim =500))
cl_window_norm_smooth <- lapply(cl_window_norm_smooth, function(x) clean_smooth(data_smoothed = x))  


# heatmap of all three
#cl_window_norm_smooth <- rbind(cl_window_norm_smooth$endo, cl_window_norm_smooth$oe, cl_window_norm_smooth$flag)
h1 <- Heatmap(cl_window_norm_smooth[[1]], cluster_columns = F, use_raster = T, raster_device = "png", raster_quality = 10) 
h2 <- Heatmap(cl_window_norm_smooth[[2]], cluster_columns = F, use_raster = T, raster_device = "png", raster_quality = 10) 
h3 <- Heatmap(cl_window_norm_smooth[[3]], cluster_columns = F, use_raster = T, raster_device = "png", raster_quality = 10)  

### TODO MB
### resulting heatmaps look a bit strange, maybe try different sorting of ranges
### be aware that clustering is only performed on h1 and all heatmaps get there 
### rows arranged by that clustering
###

h1+h2+h3



```

## in the beginning of CDSs

```{r eval=FALSE, include=TRUE}
### TODO MB
### same as for CDS
###

### ============================================
###  Crosslink patterns at the start of CDS
### ============================================

# window at the beginning of the 3'UTR
# select UTRs
CDS <- annotation[annotation$type=="CDS"] 
CDS <- CDS[!duplicated(three_UTRs$gene_id)]
w = 300

window <- CDS
end(window[strand(window)=="+"]) <- start(window[strand(window)=="+"])+300
start(window[strand(window)=="-"]) <- end(window[strand(window)=="-"])-300


cl_window <- c(map(hela_cl_rle[c(1,3,5)], ~.x[window[strand(window)=="+"]] %>% as.matrix() ),
               map(hela_cl_rle[c(2,4,6)], ~.x[window[strand(window)=="-"]] %>% as.matrix() %>% .[,ncol(.):1] ))

# select 3'UTRs with medium coverage
cl_window_coverage_p <- map_dfc(cl_window[1:3], ~ rowSums(.x))
cl_window_coverage_m <- map_dfc(cl_window[4:6], ~ rowSums(.x))

cov_p <- map_dfc(cl_window_coverage_p, ~.x) %>% mutate(medium_cov =
                                                            case_when( endo_p > 100 &
                                                                        oe_pura_p > 100 &
                                                                        oe_flag_p > 100 &
                                                                        endo_p < 10^6 &
                                                                        oe_pura_p < 10^6 &
                                                                        oe_flag_p < 10^6 
                                                                      ~ T,
                                                                      T~F))
cov_m <- map_dfc(cl_window_coverage_m, ~.x) %>% mutate(medium_cov =
                                                            case_when( endo_m > 100 &
                                                                        oe_pura_m > 100 &
                                                                        oe_flag_m > 100 &
                                                                        endo_m < 10^6 &
                                                                        oe_pura_m < 10^6 &
                                                                        oe_flag_m < 10^6 
                                                                      ~ T,
                                                                      T~F))


cl_window <- c(map(cl_window[1:3], ~.x[cov_p$medium_cov,] ),
               map(cl_window[4:6], ~.x[cov_m$medium_cov,] ))


window_p <- window[strand(window)=="+"]
window_m <- window[strand(window)=="-"]
window <- c(window_p[cov_p$medium_cov], window_m[cov_m$medium_cov])


cl_window <- list(endo = rbind(cl_window[[2]], cl_window[[5]]),
                  oe  = rbind(cl_window[[1]], cl_window[[4]]),
                  flag = rbind(cl_window[[3]], cl_window[[6]]))


# normalise
cl_window_normalized <- lapply(cl_window, function(x) apply(x, 1, min_max_noramlize) %>% t(.))


# set NaNs to zero
cl_window_normalized[[1]][which(is.na(cl_window_normalized[[1]]))] = 0
cl_window_normalized[[2]][which(is.na(cl_window_normalized[[2]]))] = 0
cl_window_normalized[[3]][which(is.na(cl_window_normalized[[3]]))] = 0

# smoothing
cl_window_norm_smooth <- lapply(cl_window_normalized,  apply, 1, function(x) smoothing(y=x, lambda = 0.5, dim =500))
cl_window_norm_smooth <- lapply(cl_window_norm_smooth, function(x) clean_smooth(data_smoothed = x))  


# heatmap of all three
#cl_window_norm_smooth <- rbind(cl_window_norm_smooth$endo, cl_window_norm_smooth$oe, cl_window_norm_smooth$flag)
h1 <- Heatmap(cl_window_norm_smooth[[1]], cluster_columns = F, use_raster = T, raster_device = "png", raster_quality = 10) 
h2 <- Heatmap(cl_window_norm_smooth[[2]], cluster_columns = F, use_raster = T, raster_device = "png", raster_quality = 10) 
h3 <- Heatmap(cl_window_norm_smooth[[3]], cluster_columns = F, use_raster = T, raster_device = "png", raster_quality = 10)  

h1+h2+h3



```
